\input cheatmac-plain

\def\Pee{{\bb P}}
\def\Zet{{\bb Z}}
\def\Rko{{\bb R}}
\def\Bigpi{{\rm Par}}
\def\intcone{{\rm intcone}}
\def\cone{{\rm cone}}
\def\conv{{\rm conv}}
\def\vert{{\rm vert}}
\def\Oh{{\rm O}}
\def\supp{{\rm supp}}
\def\enc{{\rm enc}}
\let\cfix=\cdot

\title{Polynomiality for Bin Packing with}
\title{a Constant Number of Item Types}
\author{Michel X. Goemans, Thomas Rothvoß}
\presenter{Martin Böhm}
\centerline{{\it Combinatorics and Graph Theory PhD Seminar, 2014, MFF UK}}

%TODO: \Pee, \Zet, \Rko

\section{Bin packing}

\dfn[{\csc Bin Packing}]{

{\bf Input:} A pair of vectors $(s,a)$, where $s_1, s_2, s_3, … s_d$ are
{\it item types}, i.e. all possible sizes of our input items ($s_i ∈ [0,1]$)
and $a_1, a_2, …, a_d$ are {\it item multiplicities}, i.e. how many
items of each item type we need to pack ($a_i ∈ \Zet_{≥0}$).

{\bf Goal:} Find a minimum number of {\it bins} of capacity 1 such that all items
are packed.
}

We are only considering a constant number of item types $d$.

We can look at {\csc Bin Packing} also in this manner:

{\bf Input:} A pair of vectors $(s,a)$ as before. From these two vectors, define
a {\it configuration space} $\Pee ≡ \{x ∈ \Zet_{≥0}^d | s^T x ≤ 1 \}$. An element $x$ in
the configuration space represents one valid packing of a bin.

{\bf Goal:} Select a minimum number of vectors in $\Pee$ such that we
use all items with respect to their multiplicities, i.e. the vectors
of configuration space we use sum up to $a$:

$$\min \left\{ ∑_iλ_i | ∑_{x ∈ \Pee} λ_x \cfix x = a; λ ∈ \Zet_{≥0}^{\Pee}\right\}.$$

{\it Note:} Even for fixed $d$, both $\Pee$ and $λ_x$ will be exponentially large.

\thm[Main result]{
For any {\csc Bin Packing} instance $(s,a)$, an optimum integral
solution can be computed in time $\Oh(\log Δ)^{2^{\Oh(Δ)}}$, where $Δ$ is the largest
integer appearing in the denominator $s_i$ or in a multiplicity $a_i$.
}

\section{The polyhedral cookbook}

\dfn{Given a set $X ⊆ \Rko^d $, we define a {\it convex cone} as $\cone(X) ≡ \{∑_{x ∈ X} λ_x x | λ_x ≥ 0 ∀x ∈ X \}$
and an {\it integer cone} as $\intcone(X) = \{∑_{x ∈ X} λ_x x | λ_x ∈ \Zet_{≥0} ∀x ∈ X \}$.
}

\dfn{
For a polytope $P = \{x ∈ \Rko^d | Ax ≤ b \}$, we define $\enc(P)$ as the number of bits
that it takes to write down the inequalities defining $P$.
}

\dfn{For a vector $λ$ we define {\it support} $\supp(λ)$ as the non-zero indices of $λ$.}

\dfn{Define a $d$-dimensional {\it parallelepiped} $Π$ with center $v_0$ as
$$ Π = \left\{v_0 + ∑_{i=1}^k μ_i v_i : |μ_i| ≤ 1 \right\}. $$
Usually we assume that parallelepipeds have linearly independent
vectors $v_i$.
}

\thm[Finding conic combinations]{Given polytopes $P,Q ⊆ \Rko^d$, one can find a $y ∈ \intcone(P ∩ \Zet^d) ∩ Q$ and
a vector $λ ∈ \Zet_{≥ 0}^{P ∩ \Zet^d}$ such that $y = ∑_{x ∈ P ∩ \Zet_d} λ_x x$ in time $\enc(P)^{2^{\Oh(d)}} \cfix \enc(Q)^{O(1)}$,
or decide that no such $y$ exists. Moreover, $|\supp(λ)|$ is upper bounded by $2^{2d+1}$.
}

The previous theorem can be proven using the Structure Theorem, stated as follows:

\thm[Structure Theorem]{ Let $P = \{x ∈ \Rko^d | Ax ≤ b \}$ be a polytope with $A ∈ \Zet^{m × d}$,
$b ∈ \Zet^m$ such that all coefficients are absolute-bounded by $Δ$. Then there exists a set
$X ⊆ P ∩ \Zet^d$ of size $|X| ≤ N ≡ m^d d^{\Oh(d)} (\log Δ)^d$ that can be computed in time $N^{\Oh(1)}$
with the following property:

For any vector $a ∈ \intcone(P ∩ \Zet^d)$ there exists an integral vector $λ ∈ \Zet_{≥ 0}^{P ∩ \Zet^d}$ such
that $∑_{x ∈ P ∩ \Zet^d} λ_x \cfix x = a$ and
\numlist\ndotted
\: $λ_x ∈ {0,1}$
\: $|supp(λ)| ≤ 2^{2d}$
\: $|supp(λ) ∖ X| ≤ 2^{2d}$.
\endlist
}

\section{The recipe}
The key idea behind the Structure Theorem is as follows:

\itemize\ibull
\: Split the polytope into polynomially many full-di\-men\-sio\-nal cells. The cells
are not equicardinal, their sizes are chosen strategically.
\: For each cell, we do the following:
\itemize\ibull
\: We fix an arbitrary integral point of the cell.
\: We envelop all integral points of the cell by a blowup convex hull with few vertices.
\: Using the hull, we cover all integral points with polynomially many parallelepipeds.
\: If too many points are selected into $λ_x$, we redistribute their weight to the vertices of the parallelepiped.
\endlist
\endlist

\section{The pre-baked ingredients}

\thm[Solving integer programs of fixed dimension]{
Given $A∈\Zet^{m × d}$ and $b ∈ \Zet^m$ with $Δ ≡ \max(||A||_∞, ||b||_∞)$,
one can find an $x ∈ \Zet^d$ with $Ax ≤ b$ (or deciding that none exists) in time $d^{\Oh(d)} \cfix m^{\Oh(1)}$.
}

\thm[Few vertices in an int. hull]{
Consider any polytope $P$ with $m$ constraints and $Δ ≡ \max(||A||_∞, ||b||_∞) ≥ 2$. 
Then $P_I = \conv(P ∩ \Zet^d)$ has at most $(m \cfix \Oh(\log Δ))^d$ extreme points. In fact
a list of the extreme points can be computed in time $d^{\Oh(d)}(m \cfix \Oh(\log Δ))^{\Oh(d)}$.
}

\thm[Encapsulate a polytope by a blowup with few vertices]{
For a centrally symmetric polytope $P ⊆ \Rko^d$, there are $k ≤ {d \over 2} (d+3)$ many
extreme points $x_1, …, x_k ∈ \vert(P)$ such that $P ⊆ \conv(± \sqrt{d} \cfix x_j | j ∈ [k])$.
}

\thm[Computing a minimum volume ellipsoid]{
Given a set of points $S$ in $\Rko^d$, we can use SDP to compute a minimum volume ellipsoid $E$
containing the given points in time polynomial to their encoding. Moreover, using the dual
solution of the SDP, we can determine the contact points of $E ∩ \conv(S)$.
}

\section{Cooking}

\lem[1]{
Let $P = \{x ∈ \Rko^d | Ax ≤ b \}$ be a polytope defined by $m$ inequalities with integral
coefficients of absolute value at most $Δ$. Then there exists a set $\Bigpi$ of at most $N ≡ m^d d^{\Oh(d)} (\log Δ)^d$
integral parallelepipeds such that
$$P ∩ \Zet^d ⊆ ⋃_{Π ∈ \Bigpi} Π ⊆ P.$$
}

\lem[2]{
For any polytope $P ⊆ \Rko^d$ and any integral vector $λ ∈ \Zet_{≥ 0}^{P ∩ \Zet^d}$ there exists
a $μ ∈ \Zet_{≥0}^{P ∩ \Zet^d}$ such that $|\supp(μ)| ≤ 2^d$ and $∑_x μ_x x = ∑_x λ_x x$. Furthermore,
$\supp(μ) ⊆ \conv(\supp(λ))$.
}

\lem[3]{
Given an integral parallelepiped $Π$ with vertices $X$. Then for any $x^* ∈ Π ∩ \Zet^d$ and
$λ^* ∈ \Zet_{d ≥ 0}$ there is an integral vector $μ ∈ \Zet_{≥ 0}^{Π ∩ \Zet^d}$ such that the following
holds:

\numlist\ndotted
\: $λ^* x^* = ∑_x μ_x x$,
\: $|\supp(μ ∖ X)| ≤ 2^d$,
\: $μ_x ∈ \{0,1\} ∀x ∈ X$.
\endlist
}

\prf[Finding conic combinations]{ Let $P = \{x | Ax ≤ b\}, Q = \{x| \overline{A}x ≤ \overline{b}\}$.

Compute the set $X$ of size at most $N = m^d d^{\Oh(d)} (\log Δ)^d$ from the Structure
Theorem in time $N^{\Oh(1)}$. Let $y^*$ be the (unknown) target vector.
Using the Structure Theorem, we get $λ^*$.

At the expense of a factor $N^{2^{2d}}$ guess $X' = X ∩ \supp(λ^*)$. At the expense
of factor $2^{2d} + 1$ guess the number $k = ∑_{x ∈ X'} λ^*_x ∈ [2^{2d}]$ of extra points.

Create the following ILP:

$$\eqalign{
∀i ∈ [k]: Ax_i &≤ b \cr
∑_{x ∈ X'} λ_x x + ∑_{i=1}^k x_i &= y \cr
\overline{A}y &≤ \overline{b} \cr
∀x ∈ X': λ_x &∈ \Zet_{≥ 0} \cr
∀i ∈ [k]: x_i &∈ \Zet^d \cr
}$$

The number of variables is $X' + (k+1)d ≤ 2^{\Oh(d)}$, the number
of constraints is $km + d + \overline{m} + |X'|d = 2^{\Oh(d)}m + \overline{m}$.
Maximal coefficient is $\max(d! Δ^d, \overline{Δ})$. 
}

{\it Bon appetit!}

\bye
